---
title: "Time Series"
author: "Teja Alluru"
date: "11/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("D:/MSBA/Autumn/TBANLT540/Time Series/Lab-2"))
```


Library Packages

```{r library}

library(tidyverse)
library(readxl)
library(corrgram)
library(tseries)
library(forecast)
```


```{r }
RainAtlantaAll <- read_csv("Atlanta_Precipitation.csv")

```

```{r data structure}
dim(RainAtlantaAll)    #returns the dimensions of an object
str(RainAtlantaAll)
sum(is.na(RainAtlantaAll)) #returns how many observations have "na"

```


```{r}
#RainAtlantaAll$SNOW <- as.character(as.logical(RainAtlantaAll$SNOW))
#RainAtlantaAll$ACMH <- as.character(as.logical(RainAtlantaAll$ACMH))
#RainAtlantaAll$ACSH <- as.character(as.logical(RainAtlantaAll$ACSH))
#RainAtlantaAll$WDFG <- as.character(as.logical(RainAtlantaAll$WDFG))
#RainAtlantaAll$WDFM <- as.character(as.logical(RainAtlantaAll$WDFM))
#RainAtlantaAll$WSF1 <- as.character(as.logical(RainAtlantaAll$WSF1))
#RainAtlantaAll$WSFG <- as.character(as.logical(RainAtlantaAll$WSFG))
#RainAtlantaAll$WSFM <- as.character(as.logical(RainAtlantaAll$WSFM))
#RainAtlantaAll$WT07 <- as.character(as.logical(RainAtlantaAll$WT07))
#RainAtlantaAll$WT09 <- as.character(as.logical(RainAtlantaAll$WT09))
#RainAtlantaAll$WT11 <- as.character(as.logical(RainAtlantaAll$WT11))
#RainAtlantaAll$WT15 <- as.character(as.logical(RainAtlantaAll$WT15))
#RainAtlantaAll$WV01 <- as.character(as.logical(RainAtlantaAll$WV01))
#RainAtlantaAll$WV20 <- as.character(as.logical(RainAtlantaAll$WV20))


#RainAtlantaAll$SNOW [is.na(RainAtlantaAll$SNOW )] <- '0' 
#RainAtlantaAll$ACMH [is.na(RainAtlantaAll$ACMH )] <- '0' 
#RainAtlantaAll$ACSH [is.na(RainAtlantaAll$ACSH )] <- '0' 
#RainAtlantaAll$WDFG [is.na(RainAtlantaAll$WDFG )] <- '0' 
#RainAtlantaAll$WDFM [is.na(RainAtlantaAll$WDFM )] <- '0' 
#RainAtlantaAll$WSF1 [is.na(RainAtlantaAll$WSF1 )] <- '0' 
#RainAtlantaAll$WSFG [is.na(RainAtlantaAll$WSFG )] <- '0' 
#RainAtlantaAll$WSFM [is.na(RainAtlantaAll$WSFM )] <- '0' 
#RainAtlantaAll$WT07 [is.na(RainAtlantaAll$WT07 )] <- '0' 
#RainAtlantaAll$WT09 [is.na(RainAtlantaAll$WT09 )] <- '0' 
#RainAtlantaAll$WT11 [is.na(RainAtlantaAll$WT11 )] <- '0' 
#RainAtlantaAll$WT15 [is.na(RainAtlantaAll$WT15 )] <- '0' 
#RainAtlantaAll$WV01 [is.na(RainAtlantaAll$WV01 )] <- '0' 
#RainAtlantaAll$WV20 [is.na(RainAtlantaAll$WV20 )] <- '0' 



sum(is.na(RainAtlantaAll))

#RainAtlantaAll[is.na(RainAtlantaAll)] <- 0 #replaces "na" with 0. This is a choice, statistically, but you can't run the regression without it

#sum(is.na(RainAtlantaAll))
summary(RainAtlantaAll)

```

Create a time series dataset for decomposition

```{r time_series_creation}

RainAtlantaAll$DATE<-as.POSIXct(RainAtlantaAll$DATE, format="%m/%d/%Y")
RainAtlantaAll$PRCP<-as.numeric(RainAtlantaAll$PRCP)

MonthlyRain<-aggregate(list(rain = RainAtlantaAll$PRCP), 
          list(month = cut(RainAtlantaAll$DATE, "month")), 
          mean)
MonthlyRain

MonthlyRain2<-ts(MonthlyRain$rain, frequency = 12, start = c(1945,1))
MonthlyRain2

Rain<-ts(RainAtlantaAll$PRCP, frequency = 365, start = c(1945,1))

```


Create a plot of the time series


```{r plots}
plot.ts(Rain)
plot.ts(MonthlyRain2)

```


Identify the trend/season/random components

```{r decomposition}
RainParts<-decompose(Rain)
RainMonthParts<-decompose(MonthlyRain2)
plot(RainParts)
plot(RainMonthParts)
```

Modeling using exponential smoothing - Full data

```{r HoltWintersFull}
RainModel1<-HoltWinters(Rain)
RainModel1
RainModel1$SSE
plot(RainModel1, col=4, col.predicted=6)
residualsHolt1<-residuals(RainModel1)
plot(residualsHolt1)
acf(residualsHolt1)
pacf(residualsHolt1)
```

Modeling using exponential smoothing - Monthly data

```{r HoltWintersMonthly}
RainModel2<-HoltWinters(MonthlyRain2)
RainModel2
RainModel2$SSE
plot(RainModel2, col=4, col.predicted=6)
residualsHolt2<-residuals(RainModel2)
plot(residualsHolt2)
acf(residualsHolt2)
pacf(residualsHolt2)
```


Forecasting using exponential smoothing - Full Data

```{r HWForecastFull}
RainForecast<-forecast(Rain, h=400)
plot(RainForecast)
```

Forecasting using exponential smoothing - Monthly Data

```{r HWForecastFull}
RainForecast2<-forecast(MonthlyRain2, h=13)
plot(RainForecast2)
```

Modeling using an auto.arima model - Full Data 

```{r AutoArimaFull}
par(mfrow=c(1,2))
acf(Rain)
pacf(Rain)


RainArima<-auto.arima(Rain, seasonal = TRUE)
RainArima
acf(ts(RainArima$residuals), main='ACF Residual - Full')
pacf(ts(RainArima$residuals), main='PACF Residual - Full')


```


Modeling using an auto.arima model - Monthly Data 

```{r AutoArimaMonthly}
acf(MonthlyRain2)
pacf(MonthlyRain2)

RainArima2<-auto.arima(MonthlyRain2, seasonal = TRUE)
RainArima2

acf(ts(RainArima2$residuals), main='ACF Residual - Monthly')
pacf(ts(RainArima2$residuals), main='PACF Residual- Monthly')

```


```{r predictions}
RainArima3<-Arima(MonthlyRain2, order = c(5, 1, 0), seasonal = c(5, 0, 0),
                       include.mean = FALSE, include.drift = FALSE, include.constant =TRUE,
                       method = "ML")

RainArima4<-Arima(MonthlyRain2, order = c(5, 1, 0), seasonal = c(4, 0, 0),
                  include.mean = FALSE, include.drift = FALSE, include.constant =TRUE,
                  method = "ML")
RainArima4



prediction1=predict(RainArima,n.ahead=10)
prediction1
plot(forecast(RainArima,h=100))

prediction2=predict(RainArima2,n.ahead=10)
prediction2
plot(forecast(RainArima2,h=100))




```


